FoRTE baseline

### 2019-07-16 analysis/paper/paper.Rmd
Minor tweaks to methods

## ED2 parameters

The full list of parameters in ED2 is large, with over 100 parameters per plant functional type.
A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
Instead, we selected a subset of parameters that were identified as important by previous ED2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].

@@

For each plant functional type, we generated a distribution of parameter values via a trait-meta analysis [@lebauer_2013_facilitating; see also @dietze_2014_quantitative; @raczka_2018_what].
Prior distributions for this meta-analysis were the same as those used in @dietze_2014_quantitative, @raczka_2018_what, and @viskari_2019_influence.
Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from Shiklomanov (2019 {dissertation, chapter 3}).
The resulting parameter distributions are shown in Figure X.
<!-- TODO more on this? -->
All parameters not described in this section were set to their ED2 PFT-specific defaults.

@@

For each factorial combination of model configurations/submodels (described above), we ran 100 ensemble members from 1901 to 2000.
Each ensemble member was initialized from ED2's "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
Driving meteorological data was 6-hourly CRU-NCEP <!-- TODO ref. More here? --> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric].
Soil texture was set to 92% sand, 7% silt, and 1% clay, per site-level observations in [@gough_2010_wood].
The initial soil moisture profile was set to the average soil moisture profile reported in the UMBS Ameriflux ancillary data (https://ameriflux.lbl.gov/sites/siteinfo/US-UMd).
The entire model execution workflow was conducted using the Predictive Ecosystem Analyzer (PEcAn, www.pecanproject.org).

@@

```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 8)}
bigfacet_cap <- paste(
  htmlfig(),
  "ED2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
)
readd(lai_pft_plot)
```

@@

Model sensitivity to specific parameters ("elasticity" _sensu_ ...) varied depending on the model configuration and the target output variable (Figure 5).
For predicting C fluxes, biomass, and leaf area, the three most important parameters in terms of model sensitivity were specific leaf area (SLA) for early and late hardwoods, leaf C:N ratio (`c2n_leaf`) for early and mid hardwoods, and V~cmax~ for mid and late hardwoods.
Both the magnitude and direction of sensitivity to these parameters varied by model configuration.
For example, in the default ED2 configuration (closed canopy, two-stream radiative transfer, static traits), net primary productivity increased with early hardwood specific leaf area, but when this configuration was altered so SLA and V~cmax~ could vary with light level, net primary productivity decreased with early hardwood specific leaf area.
Similarly, the model became much more sensitive to early- and mid-hardwood leaf C:N ratio when switching from a closed canopy to a finite canopy configuration.

>>>>

## ED-2.2 parameters

The full list of parameters in ED-2.2 is large, with over 100 parameters per plant functional type.
A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
Instead, we selected a subset of parameters that were identified as important by previous ED2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].

@@

For each plant functional type, we generated a distribution of parameter values via a trait-meta analysis [@lebauer_2013_facilitating; see also @dietze_2014_quantitative; @raczka_2018_what].
Prior distributions for this meta-analysis were the same as those used in @dietze_2014_quantitative, @raczka_2018_what, and @viskari_2019_influence.
Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from Shiklomanov (2019 {dissertation, chapter 3}).
The resulting parameter distributions are shown in Figure X.
<!-- TODO more on this? -->
All parameters not described in this section were set to their ED-2.2 PFT-specific default values.

@@

For each factorial combination of ED-2.2 configurations (described above), we ran 100 ensemble members from 1901 to 2000.
Each ensemble member was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
Driving meteorological data was 6-hourly CRU-NCEP <!-- TODO ref. More here? --> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric].
Soil texture was set to 92% sand, 7% silt, and 1% clay, per site-level observations in [@gough_2010_wood].
The initial soil moisture profile was set to the average soil moisture profile reported in the UMBS Ameriflux ancillary data (https://ameriflux.lbl.gov/sites/siteinfo/US-UMd).
The entire model execution workflow was conducted using the Predictive Ecosystem Analyzer (PEcAn, www.pecanproject.org).

@@

```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 8)}
bigfacet_cap <- paste(
  htmlfig(),
  "ED-2.2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
)
readd(lai_pft_plot)
```

@@

Model sensitivity to specific parameters ("elasticity" _sensu_ ...) varied depending on the model configuration and the target output variable (Figure 5).
For predicting C fluxes, biomass, and leaf area, the three most important parameters in terms of model sensitivity were specific leaf area (SLA) for early and late hardwoods, leaf C:N ratio (`c2n_leaf`) for early and mid hardwoods, and V~cmax~ for mid and late hardwoods.
Both the magnitude and direction of sensitivity to these parameters varied by model configuration.
For example, in the default ED-2.2 configuration (closed canopy, two-stream radiative transfer, static traits), net primary productivity increased with early hardwood specific leaf area, but when this configuration was altered so SLA and V~cmax~ could vary with light level, net primary productivity decreased with early hardwood specific leaf area.
Similarly, the model became much more sensitive to early- and mid-hardwood leaf C:N ratio when switching from a closed canopy to a finite canopy configuration.
